What do they see? Testing perceptual accuracy in a U.S. general population survey using stacked bar charts

### 2023-10-27 SDSS-2023/testing-charts-accuracy.tex
keywords
removing warning message and italicizing keywords
tex file overhauled and rendered as pdf

\usepackage{booktabs}
\usepackage{caption}
\usepackage{longtable}
\usepackage{lscape}
\author[1]{Kiegan Rice\thanks{Corresponding author. Email: rice-kiegan@norc.org}}
\author[2]{Heike Hofmann\footnote{Email: hofmann@iastate.edu}}
\author[1]{Nola du Toit\footnote{Email: dutoit-nola@norc.org}}
\author[1]{Edward Mulrow\footnote{Email: mulrow-edward@norc.org}}
\affil[1]{NORC at the University of Chicago}
\affil[2]{Department of Statistics, Iowa State University}
\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Testing Charts: viewers' perceptual accuracy in surveys},
  pdfkeywords={graphical perception, data visualization, survey data},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{Testing Charts: viewers' perceptual accuracy in surveys}
\date{}

\begin{document}
\maketitle
\begin{abstract}
The use of visuals is a key component in scientific communication.
Decisions about the design of a data visualization should be informed by
what design elements best support the audience's ability to perceive and
understand the components of the data visualization. We build on the
foundations of Cleveland and McGill's work in graphical perception,
employing a large, nationally-representative, probability-based panel of
survey respondents to test perception in statistical charts. Our
findings provide actionable guidance for data visualization
practitioners to employ in their work.
\end{abstract}
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, breakable, interior hidden, borderline west={3pt}{0pt}{shadecolor}, enhanced, sharp corners, boxrule=0pt]}{\end{tcolorbox}}\fi

  \newcommand{\hh}[1]{{\textcolor{orange}{#1}}} 
  \newcommand{\kr}[1]{{\textcolor{teal}{#1}}}.  
  \newcommand{\ejm}[1]{{\textcolor{ForestGreen}{#1}}}.  
  \newcommand{\ndt}[1]{{\textcolor{purple}{#1}}}.

@@

The viewer's employment of comparisons of the components is a crucial
step in the process of interacting with and understanding the chart.
Cleveland and McGill (1984) observed as such, and in their seminal study
defined the better visual among a pair as the one that allows viewers to
make more accurate comparisons. Based on mappings of quantitative
variables to different graphical elements, Cleveland and McGill's study
resulted in a ranking of perceptual tasks (such as identifying the
larger of two lengths or areas) from most accurate to least accurate,
which was then extended by Mackinlay (1986) to a theoretical framework
ranking tasks' order along their ordinal and nominal scales.

Cleveland and McGill's work -- while a foundational user study in
graphical perception -- utilized a small convenience sample, consisting
of only a few individuals recruited from among the authors' coworkers
and their spouses. Heer and Bostock (2010) reproduced Cleveland and
McGill's rankings using a larger sample from a crowd sourcing platform,
employing a total of 82 Amazon Mechanical Turk (mTurk) workers for the
study. Crowd sourced samples were shown by Borgo et al. (2017) to be
biased towards more male, younger, and relatively higher education
relative to the adult U.S. population as a whole; generally, convenience
and opt-in samples are not representative of the general population, a
common target audience for data visualization and scientific
communication work. Further, the populations' emphasis on higher
education individuals also leads to results which hold for groups of
individuals who may be more likely to have prior exposure to data
visualization in the context of scientific communication, or more
exposure to data topics in higher education, but may not hold across
other groups within the population.

Here, we seek to answer whether it is possible to reproduce some of the
previous findings in the context of a survey with a large,
nationally-representative set of respondents. Specifically, we present
viewers with structural variations of bar charts and ask them to answer
questions comparing the size of elements within those charts. We employ
a probability-based survey panel and run a series of perception tests
with nationally-representative samples of respondents from that panel.
The advantage of using a probability-based approach is two-fold. First,
we have access to a large sample of survey participants and thus have
greater power in making inference about graphical perceptional
abilities. Second, the sample is representative of the general adult
public in the U.S., and this allows us to test whether prior results
from convenience samples hold with a nationally representative sample
and whether there are differences in those results across demographic
subgroups. Within that context we focus on the following research
questions:

@@

We ask participants to determine which of two just-noticeably different
marked pieces (tiles) is larger within each chart, and focus on three
main sets of structural variation in the design of that chart:

\begin{itemize}
\item
  First, we vary the alignment of the pieces in question. Viewers are
  presented with two marked pieces in a chart that do not share a common
  baseline (aligned pieces), then two pieces that do share a common
  baseline (unaligned pieces).
\item
  Second, we vary the orientation of the chart -- we rotate the vertical
  stacked bar chart, and present a horizontally oriented version of the
  same chart, with identically sized marked pieces.
\item
  Finally, we change the aspect ratio of the chart and present a wider
  version of the horizontally oriented chart which has longer, but
  thinner, marked pieces.
\end{itemize}

@@

{\centering \includegraphics[width=0.9\textwidth,height=\textheight]{images/stimuli-overview.png}

@@

\caption{\label{fig-tasks}Each of the visual stimuli presented to
viewers. In each bar chart, two pieces are marked. The larger piece in
each chart are pieces B and D.}

@@

``There are many charts used in the news media to portray data visually.
Looking at the chart below, which of the marked dark blue pieces is
bigger, A or B? Just your best guess is fine.

\begin{itemize}
\item
  A is bigger
\item
  B is bigger
\item
  They are the same''
\end{itemize}

@@

Aligned tiles are additionally anchored in the same position along the
axis, i.e.~the difference between their sizes can be reduced to a
positional assessment, which in the absence of three-dimensional cues
seems to be the general strategy viewers use when evaluating barcharts
(Zacks et al. 1998). Unaligned tiles do not share this anchor of
position, i.e.~the difference between their sizes needs at least two
comparisons of positions. However, unaligned bars in a stacked barchart
are not just assessments of the lengths of bars -- the other tiles in
the chart provide a reference frame of the shared axis, which
\emph{should} help with an assessment of the tiles' sizes. As shown in
the sketch of Figure~\ref{fig-tasks-explain}, this framing leads to a
stimulus that is between Cleveland and McGill's task of `position along
the same but unaligned axis' and an assessment of length.

We would expect that comparing unaligned tiles is a harder task (with
correspondingly lower levels of accuracy) than a comparison of aligned
tiles, with the framing given by the other tiles in the same column
mitigating some of this difficulty. In particular, the additional
context of the other tiles should \emph{help} rather than \emph{hurt}
the comparison, unlike other situations, where additional information
besides the pieces involved in the comparison were found to be
distractors (Talbot, Setlur, and Anand 2014).

@@

When combining responses from different surveys, weights are rescaled
with respect to the total population, so that their order after
combining still reflects their relative importance. We \emph{combine}
(rather than cumulate) a set of \(\ell\) surveys \(S_1\), \(S_2\),
\ldots{} \(S_\ell\) (with \$ \ell \ge 2\$ ) as described in
O'Muircheartaigh and Pedlow (2002), by multiplying weights in \(S_i\) by
\(\lambda_i\), for \(1 \le i \le \ell\). \(\lambda_i \in [0,1]\) with
\(\sum_i \lambda_i = 1\) is given as \begin{equation}
\lambda_i = \frac{n_i/d_i}{\sum_{j=1}^{\ell}n_j/d_j},
\end{equation} where \(n_i\) is the nominal sample of survey \(S_i\) and
\(d_i\) are the design effects for the estimators. Here, \(d_i\) are
estimated as \begin{equation}
d_i = 1 + CV(w \in S_i)^2
\end{equation} where \(CV\) is the coefficient of variation of the
weights \(w\) within each sample, and is estimated as in Kish (1965) :
\begin{equation}
CV(w \in S) = \frac{\widehat{Var(w)}}{\bar{w}^2}.
\end{equation}

@@

\hypertarget{tbl-rounds}{}
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1667}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1667}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1667}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1667}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1667}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1667}}@{}}
\caption{\label{tbl-rounds}Survey rounds: dates, number of participants
(nominal sample size), effective sample size, sum of weights, and
factors for the combining of surveys as discussed above.}\tabularnewline
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Name
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Date
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
\# Participants
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
Effective Sample Size
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
Sum of weights \(\sum_i w_i\)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
\(\lambda_i\)
\end{minipage} \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Name
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Date
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
\# Participants
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
Effective Sample Size
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
Sum of weights \(\sum_i w_i\)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
\(\lambda_i\)
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Round 1 & April 2022 & 933 & 521.1 & 934.9 & \(\lambda_1\) = 0.343 \\
Round 2 & May 2022 & 953 & 485.7 & 953.4 & \(\lambda_2\) = 0.320 \\
Round 3 & June 2022 & 921 & 513.5 & 923.1 & \(\lambda_3\) = 0.338 \\
\textbf{Total} & --- & 2807 & 1520.8 & & \\
\end{longtable}

@@

\caption{\label{fig-rounds-demographics}Demographics of respondents in
each of the tasks. The dots show percentages of the national
demographics as estimated by the American Community Survery (ACS) 2021.}

@@

\caption{\label{fig-response-123}Responses by task across the designs.
The overlaid rectangles represent 95\% confidence intervals. The letters
in blue and orange encode significances between pairwise proportions:
two bars have a significantly different proportion (at a 5\%
significance level) if they do not share a letter. There is no
significant difference between the designs for wrong responses.}

@@

where \(\ell \in \{1, 2\}\); \(i \in \{1, 2\}\) is the comparison type
(1 = Aligned, 2 = Unaligned), and \(j \in \{1, 2, 3\}\) is the chart
design, with 1 = Vertical, 2 = Horizontal, and 3 = Horizontal wide. The
estimated values and 95\% confidence intervals are shown in
Table~\ref{tbl-rounds-123}. The lowercase letters indicate significances
between pairs of estimates: two estimates within the same column are
significantly different at 5\% level if they do not have any letter in
common (Piepho 2004).

@@

One contributing factor to the observed patterns in response accuracy
might be the way that participants interact with the different designs.
Across all tasks, about half of all participants make use of the option
to zoom into charts. We observe that while zooming does help with the
overall accuracy (which is in agreement with the findings by Lu et al.
(2022) about the physical size of stimuli), the increase is not
significant. However, different designs lead to different rates of
zooming: we observe in \autoref{fig-zoom-123} that when dealing with the
vertical design, the rate of zooming is significantly higher than for
the two horizontal designs.

@@

{\centering \includegraphics[width=0.8\textwidth,height=\textheight]{./figures/fig-zoom-123-1.png}

@@

\caption*{
{\large Estimates for logistic regression on zooming behavior}
} \\ 
\toprule
term & estimate & SE & t-statistic & p-value \\ 
\midrule
$\hat{\mu}$ & $-0.35$ & $0.15$ & $-2.4$ & 0.0155 \\ 
\midrule
\multicolumn{5}{l}{Response} \\ 
\midrule
They are the same, $\widehat{\rho_2}$ & $-0.17$ & $0.09$ & $-1.9$ & 0.0578 \\ 
Wrong, $\widehat{\rho_3}$ & $-0.06$ & $0.13$ & $-0.5$ & 0.6466 \\ 
\midrule
\multicolumn{5}{l}{Task} \\ 
\midrule
Aligned, $\widehat{\tau_2}$ & $0.00$ & $0.06$ & $-0.1$ & 0.9435 \\ 
\midrule
\multicolumn{5}{l}{Chart design} \\ 
\midrule
Horizontal wide, $\widehat{\delta_2}$ & $0.27$ & $0.15$ & $1.8$ & 0.0752 \\ 
Vertical, $\widehat{\delta_3}$ & $0.98$ & $0.13$ & $7.7$ & < 0.0001 \\ 
\bottomrule
\end{longtable}

@@

{\centering \includegraphics[width=1\textwidth,height=\textheight]{./figures/fig-certainty-1.png}

@@

Let \(Y_k\) be the response of participant \(k\), on a scale from 1 =
`wrong', 2 = `they are the same' to 3 = `correct'. We use a generalized
cumulative logistic regression, where \(\mu_i\) are intercepts
\(1 \le i < 3\), \(X_k\) are demographics of the \(k\)th participant (in
form of the model matrix), and \(\beta_{i}\) are the coefficients.

@@

\begin{verbatim}
Warning in RET$pfunction("adjusted", ...): Completion with error > abseps
\end{verbatim}

@@

With our survey framework, we have shown that we can rank structural
variations on design and tasks by their relative accuracy; we identified
significantly different rates of respondent accuracy under different
designs presenting the same data. The size of our sample aids in
identifying this signal.

Further, we can also study respondent behavior and how viewers interact
with the chart. Paradata on viewer interaction with the chart (e.g.,
zooming and time spent on each task) can be collected as well as asking
directly asking respondents questions about their certainty. Not only
can we collect this information, but we can also identify significant
differences in this behavior across different designs. Understanding
viewer interaction and engagement is a key component to designing
effective data visualizations. If a particular design leads viewers to
zoom in more to investigate it when determining a response, we might
consider that design to be more frustrating to viewers `in the wild'
when viewing it outside of the context of our tests.

Perhaps most salient among our results are the patterns in response
behavior across demographic groups. The large size of our sample, paired
with the probability-based survey panel approach, afford us the ability
to dive into response patterns among demographics across tasks and
identify significant differences among them. In particular, the
differences across educational attainment groups and income groups
underscore the importance of utilizing a representative population when
testing perception; prior results in this area may be absorbing bias in
responses due to the larger representation of individuals with higher
education levels among those study respondents.

>>>>

\usepackage{booktabs}
\usepackage{caption}
\usepackage{longtable}
\usepackage{lscape}
\author[1]{Kiegan Rice\thanks{Corresponding author. Email: rice-kiegan@norc.org}}
\author[2]{Heike Hofmann\footnote{Email: hofmann@iastate.edu}}
\author[1]{Nola du Toit\footnote{Email: dutoit-nola@norc.org}}
\author[1]{Edward Mulrow\footnote{Email: mulrow-edward@norc.org}}
\affil[1]{NORC at the University of Chicago}
\affil[2]{Department of Statistics, Iowa State University}
\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={What do they see? Testing perceptual accuracy in a U.S. general population survey using stacked bar charts},
  pdfkeywords={graphical perception, data visualization, survey data},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title[Testing stacked bar charts in surveys]{What do they see? Testing perceptual accuracy in a U.S. general population survey using stacked bar charts}
\date{}

\begin{document}
\maketitle
\begin{abstract}
The use of visuals is a key component in scientific communication.
Decisions about the design of a data visualization should be informed by
what design elements best support the audience's ability to perceive and
understand the components of the data visualization. We build on the
foundations of Cleveland and McGill's work in graphical perception,
employing a large, nationally-representative, probability-based panel of
survey respondents to test perception in stacked bar charts. Our
findings provide actionable guidance for data visualization
practitioners to employ in their work.
\end{abstract}
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, enhanced, boxrule=0pt, sharp corners, breakable, interior hidden, borderline west={3pt}{0pt}{shadecolor}]}{\end{tcolorbox}}\fi

@@

The comparisons of components is a crucial step in the process of
interacting with and understanding the chart. Cleveland and McGill
(1984) observed as such, and in their seminal study defined the better
visual among a pair as the one that allows viewers to make more accurate
comparisons. Based on mappings of quantitative variables to different
graphical elements, Cleveland and McGill's study resulted in a ranking
of perceptual tasks (such as identifying the larger of two lengths or
areas) from most accurate to least accurate, which was then extended by
Mackinlay (1986) to a theoretical framework ranking tasks' order along
their ordinal and nominal scales.

Cleveland and McGill's work -- while a foundational user study in
graphical perception -- utilized a small convenience sample, consisting
of only a few individuals recruited from among the authors' coworkers
and their spouses. Heer and Bostock (2010) reproduced Cleveland and
McGill's rankings using a larger sample from a crowd sourcing platform,
employing a total of 82 Amazon Mechanical Turk (mTurk) workers for the
study. Crowd sourced samples were shown by Borgo et al. (2017) to be
biased towards more male, younger, and relatively well-educated
individuals relative to the adult U.S. population as a whole; generally,
convenience and opt-in samples are not representative of the general
population. Further, higher education individuals are likely to have
prior exposure to data visualization or data topics in higher education,
which may not be the case for other groups within the population;
existing differences in exposure to data may drive differences in
results across groups. The general public is a common target audience
for data visualization and scientific communication work; results of
polls, surveys, and scientific work are presented often in the news,
shared on social media, or posted online or in print materials. It is
imperative that data visualizations used to communicate these topics are
designed with the goal of reaching the entire population. To that end,
an inclusive, representative approach to studying how viewers perceive
information presented in charts is critical to understand how best to
communicate with the general public.

In this work, we seek to answer whether it is possible to reproduce some
of the previous findings in the context of a survey with a large,
nationally-representative set of respondents. Specifically, we present
viewers with structural variations of bar charts and ask them to answer
questions comparing the size of elements within those charts. We employ
a probability-based survey panel and run a series of perception tests
with nationally-representative samples of respondents from that panel.
The advantage of using a probability-based approach is two-fold. First,
we have access to a large sample of survey participants and thus have
greater power in making inference about graphical perceptional
abilities. Second, the sample is representative of the general adult
public in the U.S., and this allows us to test whether prior results
from convenience samples hold with a nationally representative sample
and whether there are differences in those results across demographic
subgroups. Within that context we focus on the following research
questions:

@@

We ask participants to determine which of two just-noticeably different
marked pieces (tiles) is larger within each chart, and focus on three
main sets of structural variation in the design of that chart. First, we
vary the alignment of the pieces in question. Viewers are presented with
two marked pieces in a chart that do not share a common baseline
(aligned pieces), then two pieces that do share a common baseline
(unaligned pieces). Second, we vary the orientation of the chart -- we
rotate the vertical stacked bar chart, and present a horizontally
oriented version of the same chart, with identically sized marked
pieces. Finally, we change the aspect ratio of the chart and present a
wider version of the horizontally oriented chart which has longer, but
thinner, marked pieces.

@@

{\centering \includegraphics[width=1\textwidth,height=\textheight]{images/stimuli-overview-1.png}

@@

\caption{\label{fig-tasks}Overview of all of the visual stimuli
presented to viewers. In each bar chart, two pieces are marked. The
larger piece in each chart are pieces B and D. Please refer to the
appendix for full size images of the stimuli.}

@@

``There are many charts used in the news media to portray data visually.
Looking at the chart below, which of the marked dark blue pieces is
bigger, A or B? Just your best guess is fine. {[}{]} A is bigger; {[}{]}
B is bigger; {[}{]} They are the same.''

@@

Aligned tiles are additionally anchored in the same position along the
axis, i.e.~the difference between their sizes can be reduced to a
positional assessment, which in the absence of three-dimensional cues
seems to be the general strategy viewers use when evaluating bar charts
(Zacks et al. 1998). Unaligned tiles do not share this anchor of
position, i.e.~the difference between their sizes needs at least two
comparisons of positions. However, unaligned bars in a stacked bar chart
are not just assessments of the lengths of bars -- the other tiles in
the chart provide a reference frame of the shared axis, which
\emph{should} help with an assessment of the tiles' sizes. As shown in
the sketch of Figure~\ref{fig-tasks-explain}, this framing leads to a
stimulus that is between Cleveland and McGill's task of `position along
the same but unaligned axis' and an assessment of length.

We would expect that comparing unaligned tiles is a harder task (with
correspondingly lower levels of accuracy) than a comparison of aligned
tiles, with the framing given by the other tiles in the same column
mitigating some of this difficulty. In particular, the additional
context of the other tiles should \emph{help} rather than \emph{hurt}
the comparison, unlike other situations, where additional information
besides the pieces involved in the comparison were found to be
distractors (Talbot, Setlur, and Anand 2014). Participants first saw the
chart with the unaligned comparison, followed by the chart with the
aligned comparison, but could not return to the first comparison after
completing it. While there is some risk of participants recognizing the
subsequent tasks to be the same value comparisons, the unaligned
(harder) task was shown first to reduce the likelihood of participants
carrying over an answer from the aligned (easier) task.

@@

When combining responses from different surveys, weights are rescaled
with respect to the total population, so that their order after
combining still reflects their relative importance. We \emph{combine}
(rather than cumulate) a set of \(\ell\) surveys \(S_1\), \(S_2\),
\ldots{} \(S_\ell\) (with \(\ell \ge 2\)) as described in
O'Muircheartaigh and Pedlow (2002), by multiplying weights in \(S_i\) by
\(\lambda_i\), for \(1 \le i \le \ell\). \(\lambda_i \in [0,1]\) with
\(\sum_i \lambda_i = 1\) is given as \begin{equation}
\lambda_i = \frac{n_i/d_i}{\sum_{j=1}^{\ell}n_j/d_j},
\end{equation} where \(n_i\) is the nominal sample of survey \(S_i\) and
\(d_i\) are the design effects for the estimators. Here, \(d_i\) are
estimated as \begin{equation}
d_i = 1 + CV(w \in S_i)^2
\end{equation} where \(CV\) is the coefficient of variation of the
weights \(w\) within each sample, and is estimated as in Kish (1965) :
\begin{equation}
CV(w \in S) = \frac{\widehat{Var(w)}}{\bar{w}^2}.
\end{equation}

@@

\hypertarget{tbl-rounds}{}
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1667}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1667}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1667}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1667}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1667}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1667}}@{}}
\caption{\label{tbl-rounds}Survey rounds: dates, number of participants
(nominal sample size), effective sample size, sum of weights, and
factors for the combining of surveys as discussed above.
}\tabularnewline
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Name
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Date
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
\# Participants
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
Effective Sample Size
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
Sum of weights \(\sum_i w_i\)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
\(\lambda_i\)
\end{minipage} \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Name
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Date
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
\# Participants
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
Effective Sample Size
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
Sum of weights \(\sum_i w_i\)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
\(\lambda_i\)
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Round 1 & April 2022 & 933 & 521.1 & 934.9 & \(\lambda_1\) = 0.343 \\
Round 2 & May 2022 & 953 & 485.7 & 953.4 & \(\lambda_2\) = 0.320 \\
Round 3 & June 2022 & 921 & 513.5 & 923.1 & \(\lambda_3\) = 0.338 \\
\textbf{Combined} & --- & 2807 & 1520.8 & & \\
\end{longtable}

@@

\caption{\label{fig-rounds-demographics}Overview of participants'
demographics in each of the tasks: the colored dots show marginal
percentages for demographics across the three tasks. In comparison, the
dark lines show upper and lower 95\% confidence of these demographics as
estimated by the American Community Survery (ACS) 2021. Note that for a
lot of these intervals the estimates are so close together that they
appear as a single line.}

@@

\caption{\label{fig-response-123}Responses by task across the designs.
The overlaid rectangles represent 95\% confidence intervals. The letters
in blue and orange encode significances between pairwise proportions
(Piepho 2004): two bars have a significantly different proportion (at a
5\% significance level) if they do not have a letter in common. There is
no significant difference between the designs for wrong responses.}

@@

where \(\ell \in \{1, 2\}\); \(i \in \{1, 2\}\) is the comparison type
(1 = Aligned, 2 = Unaligned), and \(j \in \{1, 2, 3\}\) is the chart
design, with 1 = Vertical, 2 = Horizontal, and 3 = Horizontal wide. The
letters allow us to assess whether differences between two means are
significant: two estimates within the same column are significantly
different (at the 5\% level) if they do not have any letter in common.
This method is called connected letters and has been introduced by
Piepho (2004). In Figure~\ref{fig-response-123}, we see, that the
proportions of wrong responses are significantly different between
tasks, but designs within each task are not significantly different. For
correct responses, all proportions are significantly different except
for the proportion of correctly assessing the difference between
unaligned rectangles in the wide horizontal design: their label of
\texttt{ab} shows us that assessing the differences between unaligned
rectangles in the wide horizontal design is not significantly better
than in the horizontal design (labelled \texttt{a}), but also not
significantly worse than in the vertical design (labelled \texttt{b}).

The estimated values and 95\% confidence intervals are shown in
Table~\ref{tbl-rounds-123}.

@@

One contributing factor to the observed patterns in response accuracy
might be the way that participants interact with the different designs.
Across all tasks, about half of all participants make use of the option
to zoom into charts. We observe that while zooming does help with the
overall accuracy (which is in agreement with the findings by Lu et al.
(2022) about the physical size of stimuli), the increase is not
significant. However, different designs lead to different rates of
zooming: we observe in \autoref{fig-zoom-123} that when a significantly
lower rate of respondents chose to make use of zooming when answering
the horizontal designs compared to the vertical design.

@@

{\centering \includegraphics[width=0.95\textwidth,height=\textheight]{./figures/fig-zoom-123-1.png}

@@

\caption*{
{\large Estimates for logistic regression on zooming behavior}
} \\ 
\toprule
term & estimate & SE & t-statistic & p-value \\ 
\midrule
\multicolumn{5}{l}{} \\ 
\midrule
$\widehat{\mu}$ & $-0.35$ & $0.15$ & $-2.4$ & 0.0155 \\ 
\midrule
\multicolumn{5}{l}{Response} \\ 
\midrule
They are the same $\widehat{\rho}_2$ & $-0.17$ & $0.09$ & $-1.9$ & 0.0578 \\ 
Wrong $\widehat{\rho}_3$ & $-0.06$ & $0.13$ & $-0.5$ & 0.6466 \\ 
\midrule
\multicolumn{5}{l}{Task} \\ 
\midrule
Aligned $\widehat{\tau}_2$ & $0.00$ & $0.06$ & $-0.1$ & 0.9435 \\ 
\midrule
\multicolumn{5}{l}{Chart design} \\ 
\midrule
Horizontal wide $\widehat{\delta}_2$ & $0.27$ & $0.15$ & $1.8$ & 0.0752 \\ 
Vertical $\widehat{\rho}_3$ & $0.98$ & $0.13$ & $7.7$ & < 0.0001 \\ 
\bottomrule
\end{longtable}

@@

{\centering \includegraphics[width=0.95\textwidth,height=\textheight]{./figures/fig-certainty-1.png}

@@

Let \(Y_k\) be the response of participant \(k\), on a scale from 1 =
`correct', 2 = `they are the same' to 3 = `wrong'.

We use a generalized cumulative logistic regression, where \(\mu_i\) are
intercepts \(1 \le i < 3\), \(X_k\) are demographics of the \(k\)th
participant (in form of the model matrix), and \(\beta_{i}\) are the
coefficients.

@@

@@

With our survey framework, we have shown that we can rank structural
variations on design and tasks by their relative accuracy; we identified
significantly different rates of respondent accuracy under different
designs presenting the same data. Our study sample is larger than prior
studies with a similar scope. That large sample size improves our power
to identify this signal.

Further, we can also study respondent behavior and how viewers interact
with the chart. Paradata on viewer interaction with the chart (e.g.,
zooming and time spent on each task) can be collected as well as
directly asking respondents questions about their certainty. Not only
can we collect this information, but we can also identify significant
differences in this behavior across different designs. Understanding
viewer interaction and engagement is a key component to designing
effective data visualizations. If a particular design leads viewers to
zoom in more to investigate it when determining a response, we might
consider that design to be more frustrating to viewers `in the wild'
when viewing it outside of the context of our tests.

Perhaps most salient among our results are the patterns in response
behavior across demographic groups. The large size of our sample, paired
with the probability-based survey panel approach, afford us the ability
to dive into response patterns among demographics across tasks and
identify significant differences among them. In particular, the
differences across educational attainment groups and income groups
underscore the importance of utilizing a representative population when
testing perception; prior results in this area may be absorbing bias in
responses due to the larger representation of individuals with higher
education levels among those study respondents. If studies using
non-representative or crowd-sourced samples find non-significant results
when studying elements of design, those results may not actually be
non-significant for other populations, particularly lower education or
lower income populations. Researchers should not disregard
non-significant findings in those survey studies but rather ask whether
the lack of significance would hold across population subgroups.

Our study results provide actionable insights that data visualization
practitioners can utilize in the design of stacked bar charts.
Comparisons of interest should be lined up and share a common baseline
within a chart; in an interactive setting, it may be beneficial to give
viewers the possibility of re-ordering categories to support effective
comparison across groupings. Allowing users to zoom in on images may not
impact perceptual accuracy, but it leads to higher certainty in viewers'
comparisons. Considering the target audience for visualizations and what
key findings are most important to convey should be an important step in
the process of designing a data visualization.
